Testable normative theories

### 2020-05-20 simple.tex
Impr proof

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I}, q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\ileadstoe[i]} \subseteq (\allargs × \allargs) × \Phi$, ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$, and $+$ a commutative associative internal binary operation on $\allargs$. Given $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument.

@@

\subsection{Theorem}
\begin{theorem}
	If a theory is answerable, has bounded length, does not repeat, and is operationally valid, it is valid.
\end{theorem}
\begin{proof}
	Lemma: given $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$, if $\ar[k] \nileadstost \phi$, then: $\exists \ar[k + 1] \in \gargs[\phi] \suchthat \card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.
	To prove the lemma, let us pick any $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$ such that $\ar[k] \nileadstost \phi$. 
	We know that $\exists \ar_1 \in \allargs \suchthat (\ar_1, \ar[k]) \nileadstost \phi$.

	We want to prove, first, that $(\ar_1, \ar[k]) \ileadstoe ¬\phi$. If $\phi = (\prop, \text{poss})$, that is clear from the definition of $\ileadstost$. If $\phi = (\prop, \text{sure})$,

	\commentOC{With new definition of $\ileadstost$, when $\phi = (\prop, \text{sure})$, from $\ar \nileadstost \phi$ we only obtain that $\ar \ileadstoe ¬\phi \lor …$, thus, we still need to show that $\ileadstoe (\prop, \text{sure})$ and that $\nileadstoe (¬\prop, \text{sure})$.}
	We know that $\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatse[!\phi] \ar[k]$. Therefore, by Answerability, $\exists \ar_1' \in \allargs \suchthat \ar_1' \gbeatsstar \ar[k]$, and, by Operational validity, $\exists \ar_2 \in \gargs[\phi] \suchthat \ar_2 \gbeats \ar'_1$. 
	Define $\ar[k + 1] = \ar[k] + \ar_2$. 
	By No repetition, $\ar_2 \notin c(\ar[k])$.
	Thus, $\card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.

	To prove the theorem, given $\phi \in {\gleadsto}(\allargs)$, we need to exhibit an argument $\ar$ such that $\ar \ileadstost \phi$.
	Define a finite sequence $(\ar[0], \ar[1], \ar[2], …)$, starting from $\ar[0] = {\gleadstoinv}(\phi)$, using the lemma, until the condition $\ar[k] \nileadstost \phi$ no longer holds. This condition must turn false after some finite number of steps, because $\card{c(\ar[k])}$ grows at each step, $c(\ar[k]) \subseteq \gargs$ by definition of $c(.)$, and using Bounded length. The last element of the sequence is the argument we look for.
\end{proof}
\commentOC{No use of of $\ar[k] + \ar_2 \ibeatse \ar_1$, and we use Op val only on planned attacks! The current protocol amounts to: each time the model is endangered, it has to produce one more arg. in support, and we check only at the end (after at worst finitely many steps, as $\gargs$ is bounded) if it has managed to convince. With the difference that the model is indicated which $\ar_1$ is endangering it, to guide selection of the ultimately decisive argument. Actually, if playing $\ar_2$ at some stage did not move $i$ and $\ar_1$ still attacks $\ar_0 + \ar_2$, we can say that the model missed answering that step, thus either we choose to give it a second chance (maybe it chose the wrong $\ar_2$ to counter the $\ar_1'$ it believed was relevant, or it needed both $\ar_2$ and $\ar_4$), or we give up in fear of losing our time. Actually, the protocol is that when applying Answerability, we are supposed to pick the $\ar_1'$ that $i$ just was given by another model, thus, if that’s still the same $\ar_1$ as the step before, the model has lost (it can’t map what it’s told to an argument it knows how to answer).}

>>>>

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I}, q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with $\zar \in \allargs, {\ileadstoe[i]} \subseteq (\allargs × \allargs) × \Phi$, ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$, and $+$ a commutative associative internal binary operation on $\allargs$. Given $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument.

@@

\subsection{Theorem}
\begin{theorem}
	If a theory is answerable, has bounded length, does not repeat, and is operationally valid, it is valid.
\end{theorem}
\begin{proof}
	Lemma: given $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$, if $\ar[k] \nileadstost \phi$, then: $\exists \ar[k + 1] \in \gargs[\phi] \suchthat \card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.
	To prove the lemma, let us pick any $\phi \in {\gleadsto}(\allargs)$ and $\ar[k] \in \gargs[\phi]$ such that $\ar[k] \nileadstost \phi$. 
	We know that $\exists \ar_1 \in \allargs \suchthat (\ar_1, \ar[k]) \nileadstost \phi$. 
	Therefore, by Answerability, $\exists \ar_1' \in \allargs \suchthat \ar_1' \gbeatsstar \ar[k]$, and, by Operational validity, $\exists \ar_2 \in \gargs[\phi] \suchthat \ar_2 \gbeats \ar'_1$. 
	Define $\ar[k + 1] = \ar[k] + \ar_2$. 
	By No repetition, $\ar_2 \notin c(\ar[k])$.
	Thus, $\card{c(\ar[k + 1])} ≥ \card{c(\ar[k])} + 1$.

	To prove the theorem, given $\phi \in {\gleadsto}(\allargs)$, we need to exhibit an argument $\ar$ such that $\ar \ileadstost \phi$.
	Define a finite sequence $(\ar[0], \ar[1], \ar[2], …)$, starting from $\ar[0] = {\gleadstoinv}(\phi)$, using the lemma, until the condition $\ar[k] \nileadstost \phi$ no longer holds. This condition must turn false after some finite number of steps, because $\card{c(\ar[k])}$ grows at each step, $c(\ar[k]) \subseteq \gargs$ by definition of $c(.)$, and using Bounded length. The last element of the sequence is the argument we look for.
\end{proof}
\commentOC{If playing $\ar_2$ at some stage did not move $i$ and $\ar_1$ still attacks $\ar_0 + \ar_2$, we can say that the model missed answering that step, thus either we choose to give it a second chance (maybe it chose the wrong $\ar_2$ to counter the $\ar_1'$ it believed was relevant, or it needed both $\ar_2$ and $\ar_4$), or we give up in fear of losing our time. Actually, the protocol is that when applying Answerability, we are supposed to pick the $\ar_1'$ that $i$ just was given by another model, thus, if that’s still the same $\ar_1$ as the step before, the model has lost (it can’t map what it’s told to an argument it knows how to answer).}
